---
title: "Data Transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#5.1 Introduction

###5.1.1 Prerequisites

```{r}
library(tidyverse)

library(nycflights13)

library(Lahman)
```

###5.1.2 nycflights13

```{r}
flights
```

###5.1.3 dplyr

List of verbs: filter, arrange, select, mutate, summarize

verb(database, function)

#5.2 Filter

Flights that occurred on Jan 1st

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
(jan1 <- filter(flights, month == 1, day == 1))
```

```{r}
filter(flights, month == 12, day == 25)
```

```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

```{r}

```

###5.2.1 Comparisons

==, >, >=, <, <=, !

```{r}
filter(flights, month = 1)
```

Error due to using "=" instead of "=="

```{r}
sin(pi/2)
```

```{r}
cos(pi/2) == 0
```

```{r}
near(cos(pi/2), 0)
```

###5.2.2 Logical Operators

or -----> |
and -----> &
not -----> !

Flights that departed in Nov **or** Dec

```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
(novdec <- filter(flights, month == 11 | month == 12))
```

```{r}
filter(flights, month %in% c(5, 6, 7, 8))
```

Flights that departed in Nov **and** arrived in Dec

```{r}
filter(flights, month == 11 & month == 12)
```

Flights that were not delayed (on departure or arrival) by more than two hours

```{r}
filter(flights, dep_delay <= 120 & arr_delay <= 120)
```

Or use the opposite.

```{r}
filter(flights, !(dep_delay > 120 & arr_delay > 120))
```

```{r}
filter(flights, is.na(dep_time))
filter(flights, !is.na(dep_time))
withoutNA <- filter(flights, is.na(dep_time))
withoutNA

filter(withoutNA, is.na(dep_time))
```

##### Exercises

1. Find all flights that

Had an arrival delay of two or more hours
Flew to Houston (IAH or HOU)
Were operated by United, American, or Delta
Departed in summer (July, August, and September)
Arrived more than two hours late, but didn’t leave late
Were delayed by at least an hour, but made up over 30 minutes in flight
Departed between midnight and 6am (inclusive)

```{r}
filter(flights, arr_delay >= 120)
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL")
filter(flights, month %in% c(7, 8, 9))
filter(flights, arr_delay > 120 & dep_delay <= 0)
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
filter(flights, dep_time >= 0 & dep_time <= 600)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
filter(flights, between(dep_time, 0, 600))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

8255 flights. Other missing columns include dep_delay, arr_time, arr_delay, air_time. These may be cancelled flights.

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}

```

###5.4 Select columns with select()

```{r}
select(flights, year, month, day)
select(flights, year:day)

select(flights, -(year:day))
select(flights, dep_time:time_hour)
```

```{r}
select(flights, starts_with('dep'))
select(flights, ends_with('delay'))
select(flights, contains('_'))
```

```{r}
rename(flights, tail_num = tailnum)
```
The rename function changes variable names.

```{r}
select(flights, dep_delay, arr_delay, everything())
```

###5.5 Add new variables with mutate()

```{r}
flights_narrow <- select(flights, year:day, ends_with('delay'), distance, air_time)

flights_narrow
```

```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance / air_time * 60)

flights_narrow
```

The verbs visualize changes but will not remember them. To keep the changes, they must be manually applied to something.

```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time * 60)
```

A new variable can be used again after creating it in the same code chunk but will not show up if not preset to something.

```{r}
transmute(flights, gain = dep_delay - arr_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time * 60)
```

###5.6 Grouped summaries with summarize()

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <- group_by(flights, year, month, day)

summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_carrier <- group_by(flights, carrier)

summarize(by_carrier, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_origin <- group_by(flights, origin)

summarize(by_origin, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_air_time <- group_by(flights, air_time)

summarize(by_air_time, delay = mean(dep_delay, na.rm = TRUE))
```

This has lots of points and can most likely be plotted.

#####5.6.1 Combining multiple operations with the pipe

```{r}
by_dest <- group_by(flights, dest)
```

```{r}
delays <- summarize(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delays
```

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

```{r}
delays <- filter(delays, count > 20, dest != 'HNL')

delays
```

The filter function can tidy up the miniscule points and outliers. The cleaned-up plot looks like this:

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

For cleanup:

```{r}
ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = 0.3) + 
  geom_smooth(se = FALSE) #se = FALSE - deletes the standard error band
```

```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarize(
    count = n(), 
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE),
  ) %>%
  filter(count > 20, dest != 'HNL')

ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = 0.3) + 
  geom_smooth(se = FALSE)
```

#####5.6.2 Missing values

na.rm tells us to remove the missin values.

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(delay))
```

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(delay, na.rm = TRUE))
```

```{r}

```

is.na asks the question: is the variable NA?

na.rm gives a command: remove all NAs.

```{r}
not_cancelled <- flights %>%
  filter(!is.na(arr_delay) & !is.na(dep_delay))

not_cancelled
```

#####5.6.3 Counts

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(delay = mean(arr_delay))

delays

arrange(delays, desc(delay))
```

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay),
    count = n()
  )

delays

arrange(delays, desc(delay))

ggplot(data = delays) +
  geom_point(mapping = aes(x = count, y = delay), alpha = 1/3)
```

```{r}
delays %>%
  filter(count > 25) %>%
  ggplot() + 
  geom_point(mapping = aes(x = count, y = delay), alpha = 0.2)
```

```{r}
Batting
```

How do we calculate batting average with this data set?

```{r}
Batting

batters <- Batting %>%
  group_by(playerID) %>%
  summarize(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB)
  )

batters %>%
  filter(ab > 100) %>%
  ggplot(mapping = aes(x = ab, y = ba)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}

```





